Fisher vector places: learning compact descriptors for place recognition

### 2014-11-14 cvpr15review.tex
results: major revision

   \subsection{Results}  
      For both databases (Pittsburgh 25k and Pittsburgh 55k) we compare results of our method (FV e-SVM) to two baselines: standard bag-of-visual-words baseline (BOW) and raw Fisher vector matching without learning (FV).  
%  For the Pittsburgh 25k database we also compare results to exemplar SVM learning without the proposed descriptor re-normalization (SVM raw).

      We perform experiments on several target Fisher vector dimensions $d\in\{128,512,2048\}$. For each method we measure performance using the percentage of correctly recognized queries (Recall) similarly to, e.g.,~\cite{Chen11,Knopp2010,Sattler-BMVC12}. The query is correctly localized if at least one of the top $K$ retrieved database images is within $20$ meters from the ground truth position of the query. Results are shown for different values of $K$ in tables~\ref{tab:recall} and \ref{tab:recall2}. For the Pittsburgh 25k we also show results in the form of a curve in figure~\ref{fig:recall}. The results clearly demonstrate the benefits of the learnt descriptors with respect to the standard Fisher vectors for all target dimensions and lengths of shortlist $K$. The benefits of discriminative learning are specially prominent for low-dimensional compact descriptors ($d=128$). 
      The proposed method also significantly outperforms the bag-of-visual-words baseline. Figure~\ref{fig:images} shows examples of place recognition results.

      \paragraph{Applying our learning method to other descriptors.}
        \textcolor{petr}{ 
         The proposed approach can be applied to other descriptors beyond Fisher vectors. To demonstrate this we have applied the proposed method to the bag-of-visual-words descriptor and observed improvement from 28.7\% recall@K=1 (BOW) to 31.8\% recall@K=1 (BOW e-SVM).  
For bag-of-visual-words,  the e-SVM with p-value calibration of~\cite{Gronat13} (BOW p-val) gives a slightly better results (33.6\% recall@K=1) but at an increased training cost and a significantly higher required memory footprint at query time, which makes the method hard to scale-up to database sizes beyond 200k images. The results are given in table \ref{tab:recall} (top four rows).
        }

      \paragraph{Comparison to other methods.}
         %On the Pittsburgh 25k database, we compare performance of our learnt discriminative descriptors to the methods of~\cite{Gronat13} and~\cite{Knopp2010}, who report on the same testing data top $K=1$ recall of 36.5\% and 41.9\%, respectively (results taken from~\cite{Gronat13}). Our method outperforms~\cite{Knopp2010} already for dimension $d=128$ (37.8\%) and~\cite{Gronat13} for dimension $d=512$ (47.6\%). Furthermore, note that~\cite{Gronat13,Knopp2010} are based on a bag-of-visual-words representation, which typically needs to store between 1000-2000 non-zero visual words per image, which is significantly more than our learnt 128 or 512 dimensional descriptor. 
         On the Pittsburgh 25k database, we compare performance of our learnt discriminative descriptors to the method of~\cite{Gronat13} who report results on similar testing data. Our method outperforms~\cite{Gronat13} already for dimension $d=128$ (37.8\%).
  %       \\
         We have also tested the p-val calibration of~\cite{Gronat13} on e-SVM Fisher vectors. We found that the calibration did not perform well for the learnt Fisher vector classifiers (top 1 recall of 25.3\% compared to baseline performance of 33.6\% for dimension 128). When examining the results, we believe this may be attributed to the fact that Fisher vectors are low dimensional compared to the extremely high-dimensional (d=100,000) bag-of-visual-words representation, which affects the distribution of the classifier scores on the negative test data.
%         \\
         Finally, we investigate the performance of a raw exemplar-SVM (raw e-SVM) without any further calibration for both image descriptors . We observe that while using raw SVM scores directly without any calibration fails for the bag-of-visual-words representation (BOW raw SVM) (as reported in \cite{Gronat13}), it marginally improves recall for larger $K$ for Fisher vectors (68.4\% for FV128 raw SVM compared to 67.7\% for non-learned FV128).
%         as indicated in \ref{tab:recall} (FV128 raw SVM).
          %We have compared the proposed method with the p-value calibration of (Gronat et al.[12]) applied to Fisher vectors. The results show that our method significantly outperforms the p-value calibration (37.8\% recall@1 vs. 32.3\% recall@1; 56.9\%recall@5 vs. 50.0\% recall@5) for the Fisher vector of dimension 128. Similar gains are observed for other FV dimensions.

@@

         \begin{table}[t]
            \begin{centering}
               \include{table55k}
               \caption{ \textbf{Results on Pittsburgh 55k dataset.}
                  The fraction of correctly recognized queries (recall@K)
                  \textcolor{petr}{
                  in top $K\in\{1,2,5,10,20\}$ retrieved images.
                  The proposed method (\emph{e-SVM}) has been applied to Fisher vectors compressed to different dimensions.
                  } 
                  The learnt descriptors by the proposed method (FV e-SVM) consistently improve over the raw Fisher vector descriptors across the whole range of $K$  and all dimensions.
%                   on both the 25k and 55k Pittsburgh image datasets.     
               }
            \label{tab:recall2}
            \end{centering}
         \end{table}
         %
         %
         %
         \begin{figure*}[h]
            \include{myFigureWnorm}
            \caption{
                \textbf{Examples of correctly and incorrectly localized queries for the learnt Fisher vector representation.}
                Each example shows a query image (left) together with correct (green) and incorrect (red) matches from the database obtained by the learnt Fisher vector representation \emph{e-SVM} method (top) and the standard Fisher vector baseline (bottom) for dimension 128. Note that the proposed method is able to recognize the place depicted in the query image despite changes in viewpoint, illumination and partial occlusion by other objects (trees, lamps) and buildings. 
                Note that the baseline methods often finds images depicting the same buildings but in a distance whereas our learnt representation often finds a closer view better matching the content of the query.       
            }
            \label{fig:images}
         \end{figure*}
         %
         %
         %

>>>>

   \subsection{Results} 
      \textcolor{petr}{
         The experiments are performed on two datasets (Pittsburgh 25k and Pittsburgh 55k) on several target Fisher vector dimensions $d\in\{128,512,2048\}$. The proposed approach can be also applied to other descriptors beyond Fisher vectors. To demonstrate this we have applied the proposed method to the bag-of-visual-words.
         \\
         For each method we measure performance using the percentage of correctly recognized queries (Recall) similarly to, e.g.,~\cite{Chen11,Knopp2010,Sattler-BMVC12}. The query is correctly localized if at least one of the top $K$ retrieved database images is within $20$ meters from the ground truth position of the query. Results are shown for different values of $K$ in tables~\ref{tab:recallBOW},~\ref{tab:recall} and \ref{tab:recall2}. For the Pittsburgh 25k we also show results in the form of a curve in figure~\ref{fig:recall}. The results clearly demonstrate the benefits of the learnt descriptors with respect to the standard Fisher vectors for all target dimensions and lengths of shortlist $K$. The benefits of discriminative learning are specially prominent for low-dimensional compact descriptors ($d=128$). Figure~\ref{fig:images} shows examples of place recognition results.
      }

      % For both databases (Pittsburgh 25k and Pittsburgh 55k) we compare results of our method (FV e-SVM) to two baselines: standard bag-of-visual-words baseline (BOW) and raw Fisher vector matching without learning (FV).

      % We perform experiments on several target Fisher vector dimensions $d\in\{128,512,2048\}$. For each method we measure performance using the percentage of correctly recognized queries (Recall) similarly to, e.g.,~\cite{Chen11,Knopp2010,Sattler-BMVC12}. The query is correctly localized if at least one of the top $K$ retrieved database images is within $20$ meters from the ground truth position of the query. Results are shown for different values of $K$ in tables~\ref{tab:recall} and \ref{tab:recall2}. For the Pittsburgh 25k we also show results in the form of a curve in figure~\ref{fig:recall}. The results clearly demonstrate the benefits of the learnt descriptors with respect to the standard Fisher vectors for all target dimensions and lengths of shortlist $K$. The benefits of discriminative learning are specially prominent for low-dimensional compact descriptors ($d=128$). 
      % The proposed method also significantly outperforms the bag-of-visual-words baseline. Figure~\ref{fig:images} shows examples of place recognition results.

      \paragraph{Bag-of-visual-words}
         \textcolor{petr}{
            For bag of visual words image representation we provide three kind of results summarized in table \ref{tab:recallBOW}. First, similarly to \cite{Gronat13}, we compare proposed method (e-SVM) to the standard bag-of-visual-words baseline (BOW). We observe improvement from 28.7\% recall@K=1 (BOW) to 31.8\% recall@K=1 (BOW e-SVM). Secondly, the results are also confronted to the p-value calibration of \cite{Gronat13}. The p-value calibration (BOW p-val) gives marginally better results (33.6\% recall@K=1) but at an increased training cost and a significantly higher required memory footprint at query time, which makes the method hard to scale-up to database sizes beyond 200k images. Finally, we investigate the performance of a raw exemplar-SVM (raw e-SVM) with no calibration. As reported in \cite{Gronat13}, we observe that using these scores directly with no calibration does not perform on BOW. However, this method gives surprisingly good results on Fisher vectors. Details will be given next.
         }

         \begin{table}[b]
            \begin{centering}
               \include{table25kBOW}
            \caption{ \textbf{Results on Pittsburgh 25k dataset - BOW.}
                  The fraction of correctly recognized queries (recall@K)
                  \textcolor{petr}{
                     in the top $K\in\{1,2,5,10,20\}$ retrieved images.
                     % Different methods have been applied to two types of descriptors, the BOW and Fisher vectors compressed to different dimensions. The learnt representations (BOW \emph{e-SVM} and BOW \emph{p-val}) outperform the raw BOW baseline as well as the learnt representation without calibration (BOW raw SVM).
                  } 
                  The learnt descriptors by the proposed method (BOW e-SVM) marginally improves over the raw bag-of-visual-words descriptors (BOW) across the whole range of $K$  and all dimensions.
%                   on both the 25k and 55k Pittsburgh image datasets.     
            }
            \label{tab:recallBOW}
            \end{centering}
         \end{table}

      \paragraph{Fisher vectors}

@@

         Table \ref{tab:recall} summarizes the results of proposed method that has been applied on Fisher vectors for several target dimensions $d\in\{128,512,2048\}$. The learnt image representation (FV e-SVM) significantly outperforms the Fisher vector baseline (FV). The improvements are consistent across different lengths $K$ of the shortlist for all target dimensions of the Fisher vectors.
         Figure \ref{fig:recall} shows the results in a form of curve and it clearly demonstrates the benefits of the learnt descriptors. For comprehensiveness the figure also shows the BOW baseline curve from table \ref{tab:recallBOW} shown in black.
         \\
         The p-val calibration of~\cite{Gronat13} has been tested on e-SVM Fisher vectors. We found that the calibration did not perform well for the learnt Fisher vector classifiers (top 1 recall of 25.3\% compared to baseline performance of 33.6\% for dimension 128). When examining the results, we believe this may be attributed to the fact that Fisher vectors are low dimensional compared to the extremely high-dimensional (d=100,000) bag-of-visual-words representation, which affects the distribution of the classifier scores on the negative test data.
         \\
         Next, we investigate the performance of a raw exemplar-SVM (raw e-SVM) without any further calibration for both image descriptors . We observe that while using raw SVM scores directly without any calibration fails for the bag-of-visual-words representation (BOW raw SVM) (as reported in \cite{Gronat13}), it marginally improves recall for larger $K$ for Fisher vectors (68.4\% for FV128 raw SVM compared to 67.7\% for non-learned FV128).
         \\
         Finally, we performed experiments on the Pittsburgh 55k dataset. Similarly to the former dataset, the results summarized in table \ref{tab:recall2} show consistent improvement of the proposed method (FV e-SVM) over the Fisher vector baseline (FV) across all target dimensions. The benefits of the proposed method in terms of memory efficiency will be discussed next.
         \begin{table}[t]
               \include{table55k}
               \caption{ \textbf{Results on Pittsburgh 55k dataset - FV.}
                  in top $K\in\{1,2,5,10,20\}$ retrieved images.
                  The proposed method (\emph{e-SVM}) has been applied to Fisher vectors compressed to different dimensions.
                  The learnt descriptors by the proposed method (FV e-SVM) consistently improve over the raw Fisher vector descriptors across the whole range of $K$  and all dimensions.
            \label{tab:recall2}

      % \paragraph{Applying our learning method to other descriptors.}
      %   \textcolor{petr}{ 
      %    The proposed approach can be applied to other descriptors beyond Fisher vectors. To demonstrate this we have applied the proposed method to the bag-of-visual-words descriptor and observed improvement from 28.7\% recall@K=1 (BOW) to 31.8\% recall@K=1 (BOW e-SVM).  
      %    For bag-of-visual-words,  the e-SVM with p-value calibration of~\cite{Gronat13} (BOW p-val) gives a slightly better results (33.6\% recall@K=1) but at an increased training cost and a significantly higher required memory footprint at query time, which makes the method hard to scale-up to database sizes beyond 200k images. The results are given in table \ref{tab:recall} (top four rows).
      %   }

%       \paragraph{Comparison to other methods.}
%          %On the Pittsburgh 25k database, we compare performance of our learnt discriminative descriptors to the methods of~\cite{Gronat13} and~\cite{Knopp2010}, who report on the same testing data top $K=1$ recall of 36.5\% and 41.9\%, respectively (results taken from~\cite{Gronat13}). Our method outperforms~\cite{Knopp2010} already for dimension $d=128$ (37.8\%) and~\cite{Gronat13} for dimension $d=512$ (47.6\%). Furthermore, note that~\cite{Gronat13,Knopp2010} are based on a bag-of-visual-words representation, which typically needs to store between 1000-2000 non-zero visual words per image, which is significantly more than our learnt 128 or 512 dimensional descriptor. 
%          \textcolor{petr}{
%          On the Pittsburgh 25k database, we compare performance of our learnt discriminative descriptors to the method of~\cite{Gronat13} who report results on similar testing data. Our method outperforms~\cite{Gronat13} already for dimension $d=128$ (37.8\%).
%   %       \\
%          We have also tested the p-val calibration of~\cite{Gronat13} on e-SVM Fisher vectors. We found that the calibration did not perform well for the learnt Fisher vector classifiers (top 1 recall of 25.3\% compared to baseline performance of 33.6\% for dimension 128). When examining the results, we believe this may be attributed to the fact that Fisher vectors are low dimensional compared to the extremely high-dimensional (d=100,000) bag-of-visual-words representation, which affects the distribution of the classifier scores on the negative test data.
% %         \\
%          Finally, we investigate the performance of a raw exemplar-SVM (raw e-SVM) without any further calibration for both image descriptors . We observe that while using raw SVM scores directly without any calibration fails for the bag-of-visual-words representation (BOW raw SVM) (as reported in \cite{Gronat13}), it marginally improves recall for larger $K$ for Fisher vectors (68.4\% for FV128 raw SVM compared to 67.7\% for non-learned FV128).
% %         as indicated in \ref{tab:recall} (FV128 raw SVM).
%           %We have compared the proposed method with the p-value calibration of (Gronat et al.[12]) applied to Fisher vectors. The results show that our method significantly outperforms the p-value calibration (37.8\% recall@1 vs. 32.3\% recall@1; 56.9\%recall@5 vs. 50.0\% recall@5) for the Fisher vector of dimension 128. Similar gains are observed for other FV dimensions.
%         }
         \begin{figure*}[h]
            \include{myFigureWnorm}
                \textbf{Examples of correctly and incorrectly localized queries for the learnt Fisher vector representation.}
                Each example shows a query image (left) together with correct (green) and incorrect (red) matches from the database obtained by the learnt Fisher vector representation \emph{e-SVM} method (top) and the standard Fisher vector baseline (bottom) for dimension 128. Note that the proposed method is able to recognize the place depicted in the query image despite changes in viewpoint, illumination and partial occlusion by other objects (trees, lamps) and buildings. 
                Note that the baseline methods often finds images depicting the same buildings but in a distance whereas our learnt representation often finds a closer view better matching the content of the query.       
            \label{fig:images}
         \end{figure*}
